---
title: "Appendix II: R code"
author:
- Sanne Meijering
- Hanne Oberman
- Gerbrich Ferdinands
output: pdf_document
---

# Initialization
```{r}
# Load packages
require(survey)
require(sampling)
require(dplyr)
require(MASS)

# Load society dataset
society <- readRDS("Understanding Society innovation pnel wave A.RDS")

# Remove unnecessary columns
society <- society[,c(1:5,8,11,12,14,53:60,62:65,76,77,81,87,89,92,94,95)]

# a_dvage is the average age over all waves. 
# Inspecting the data shows that -6 gives the age during the first wave.
society$a_dvage <- as.numeric(society$a_dvage) - 6
```

# Question 2
Investigation of design weights

```{r}
# Calculate variance of design weight
Var <- var(society$a_psnenip_xd)
# Combine variance with other descriptive statistics
Descr <- cbind(t(summary(society$a_psnenip_xd)), Var)
# Print with up to 2 decimals
round(Descr, 2)

# Create histogram of design weight values
hist(society$a_psnenip_xd, 
     breaks = 8,
     main = NULL, #"Histogram of design weight variable 'a_psnenip_xd'",
     xlab = "Value of the design weight")
```

# Question 3
```{r}
# Investigate levels of government office region variable
levels(society$a_gor_dv)
nrow(society[society$a_gor_dv == "missing",])
nrow(society[society$a_gor_dv == "northern ireland",])
# None of the value are missing or northern ireland, so those can be ignored

# Run linear regression
coeff <- with(society, lm(a_psnenip_xw ~ a_psnenip_xd + a_sex + a_dvage + a_gor_dv))

# Get coefficients of government office regions
coeff <- coeff$coefficients
coeff.gor <- coeff[5:15]
# Add the base, which has value 0
coeff.gor[11] <- 0
names(coeff.gor)[11] <- "a_gor_dvnorth east"

# Sort and plot coefficients
plot(sort(coeff.gor), xlab = "Government Office Region", ylab = "Coefficient")

# Show coefficients
coeff.gor
```

In accordance with the data, Scotland, London and the West Midlands form one category each, with the fourth category containing all other areas.

```{r}
# Gather grouped region in one variable
society$gor_groups <- "England_Wales"
society$gor_groups[society$a_gor_dv=="london"] <- "London"
society$gor_groups[society$a_gor_dv == "scotland"] <- "Scotland"
society$gor_groups[society$a_gor_dv == "west midlands"] <- "West Midlands"

# Change government office region groups and age into factors
society$gor_groups <- as.factor(society$gor_groups)
society$age_fac <- as.factor(society$a_dvage)

# Rerun linear regression
coeff2 <- with(society, lm(a_psnenip_xw ~ a_psnenip_xd + a_sex + gor_groups + age_fac))

# Get coefficients
coeff2 <- coeff2$coefficients
coeff2.age <- coeff2[7:102]

# Change names of coefficients into numbers
x <- unlist(strsplit(names(coeff2.age), "age_fac"))
names(coeff2.age) <- x[x!=""]

# Plot coefficients
plot(as.numeric(names(coeff2.age)), coeff2.age, xlab="Age", ylab="Coefficient")
for (i in seq(4,104,10)+0.5){
 abline(v=i, col="red") 
}

# Plot with group changes
plot(as.numeric(names(coeff2.age)), coeff2.age, xlab="Age", ylab="Coefficient")
for (i in c(14.5,24.5,34.5,44.5,64.5,74.5)){
 abline(v=i, col="red") 
}
```

In accordance with the results, age is split up in sections of 10 years, with the 0-10 and 11-20 as well as the 51-60 and 61-79 categories combined and all people of age 81 and older placed into one group.

```{r}
# Assign the levels
levels(society$age_fac)
levels(society$age_fac) <- c(rep("group0_14", 15), rep("group15_24", 10), 
                               rep("group25_34", 10), rep("group35_44",10), 
                               rep("group45_64", 20), rep("group65_74", 10), 
                               rep("group_75", 22))

# Create table
with(society, table(age_fac,gor_groups))
# None of the groups has zero observations

# Run linear regression again
coeff3 <- with(society, lm(a_psnenip_xw ~ a_psnenip_xd + a_sex + gor_groups + age_fac))

#Get coefficients
coeff3 <- coeff3$coefficients
coeff3
```

# Question 4
```{r}
# Investigate the a_employ variable.
levels(society$a_employ[society$a_dvage > 15 & society$a_dvage < 64])
# The variable a_employ has seven levels.

summary(society$a_dvage[society$a_employ=="yes"])
summary(society$a_dvage[society$a_employ=="no"])
# Yes and no contain people of over 21 years of age.

nrow(society[society$a_employ=="missing",])
nrow(society[society$a_employ=="proxy respondent",])
# Missing and proxy respondent do not appear in the data

summary(society$a_dvage[society$a_employ=="inapplicable"])
# Inapplicable seems to contain all children and youths of 21 years and younger
# It cannot be assumed that none of them is employed. It can however be assumed that only 
# a small part of them is employed, as children under 15 cannot be employed legally
# and most would still be going to a school or university.

nrow(society[society$a_employ=="refuse",])
nrow(society[society$a_employ=="don\'t know",])
# Both refuse and don't know contain one row. These can be treated as missing data.

# Thus, our goal is to compare the proportion of employed people (yes) of working age 
# against the number of unemployed people (no, inapplicable), excluding the missing 
# data (refuse, don't know)

with(society, nrow(society[a_employ=="yes" & a_dvage >64,]))
# There are people older than 64 still working, we should exclude those. 
with(society, nrow(society[a_employ=="yes" & a_dvage <15,]))
# No one younger than 15 years is reported to be working, which is to be expected as it 
# was not a question asked to people under 21 years of age.

# Since we wish to know the proportion of employed people of working age, we need 2 groups:
# one with employed adults and one with unemployed people and employed elderly.
society$employ_dv <- as.numeric(0)
society$employ_dv[society$a_employ=='yes' & society$a_dvage <= 65] <- 1

# Create design
# Don't remove the missing values yet, as the weights are calculated including missing values 
Design <- svydesign(ids=~a_hidp, strata=~a_strata, data=society, weights=~a_psnenip_xw)
# Make a subset of non-missing values
Nonmiss <- with(Design, subset(Design, a_employ!="refuse" & a_employ!="don\'t know"))
svymean(~employ_dv, Nonmiss)
confint(svymean(~employ_dv, Nonmiss))
# 46,2% of the population is employed, with a 95% confidence interval of 44.3%-48.1%
```

# Question 5b
```{r}
# Inspect levels of variables
levels(society$a_livesp_dv)
levels(society$a_cohab_dv)
levels(society$a_single_dv)
levels(society$a_mastat_dv)

###### HOUSEHOLD SIZE PER HOUSEHOLD
# a_hidp - Household identifier
# Household size: count for each household, how many persons there are in.
# Get household size per household
count <- as.matrix(table(society$a_hidp))
summary(count)
# Turn into dataframe and join to society 
households <- data.frame(a_hidp=as.numeric(rownames(count)), hh_size=count)

##### NUMBER OF CHILDREN PER HOUSEHOLD
# Count the number of kids under the age of 16 using the mastat variable
for(i in households$a_hidp){
  hh <- society[society$a_hidp==i,]
  households$n_child[households$a_hidp==i]=as.numeric(table(hh$a_mastat_dv)["Child under 16"])
}

##### ANY CHILDREN IN HOUSEHOLD
# Create variable whether the household has children in it (True for households with children)
households$with_child <- households$n_child > 0

##### ANY SINGLE ADULTS IN HOUSEHOLD
# Create variable whether the person is a single adult (true) or not (false)
society$single_adult <- society$a_single_dv == "Yes" & society$a_dvage >= 16
# Create variable whether there is a single adult in the household
for(i in households$a_hidp){
  households$hasSingle[households$a_hidp==i] <- any(society$single_adult[society$a_hidp==i])
}

##### ANY COUPLES IN HOUSEHOLD
# Create variable whether the person is in a couple (true) or not (false)
society$inacouple <- society$a_livesp_dv == "Yes" | society$a_cohab_dv == "Yes"
# Create variable whether there is a couple in the household
for(i in households$a_hidp){
  households$hasCouple[households$a_hidp==i] <- any(society$inacouple[society$a_hidp==i])
}

##### HOUSEHOLD TYPE 
# Create matrix containing household states
household_states <- matrix(c("Couple with children", "Couple without children", 
                             "Single with children", "Single without children"), nrow=2)
# Create household info variable
households$hh_type <- ""

for (i in seq(1,nrow(households))){
  # Get right row of household state
  if (households$with_child[i]){
    child = 1
  }
  else child = 2
  
  # Get right column of household state
  if(households$hasCouple[i]){
    state = 1
  }
  else if(households$hasSingle[i]){
    state = 2
  }
  # Select household state and put in households$hh_type
  households$hh_type[i] <- household_states[child, state]
}

# Show how many of each category there are
table(households$hh_type)

# Join to society dataset
society <- left_join(society, households, by="a_hidp")
```

# Question 5d
```{r}
# Create two new age variables: 
# one to plot with the five-year intervals, 
# one with categories to compute the chi square statistic
society$a_agegr5a_dv <- factor(society$a_agegr5_dv) #rename levels for readability
levels(society$a_agegr5a_dv) <- c("0-4", "5-9", "10-14", "15-19", "20-24", "25-29",
                                  "30-34", "35-39", "40-44", "45-49", "50-54", "55-59", 
                                  "60-64", "65-69", "70+") 
society$a_agegr11_dv <- factor(society$a_agegr13_dv) 
# Combine levels so no NAs exist in the contingency table
levels(society$a_agegr11_dv) <- c(rep("0-17", 2), "18-19",  "20-24",  "25-29",  "30-34",
                            "35-39", "40-44",  "45-49", "50-54", "55-59",  rep("60+",2)) 

# Update sampling design 
Design <- svydesign(ids=~pidp, strata=~a_strata, data=society, weights=~a_psnenip_xw)

# Create stacked barplot with 5 year interval variable
tab5 <- svytable(~a_agegr5a_dv + hh_type, Design) # creates a contingency table  
plot(tab5, xlab = "Age group (years old)", ylab = "Household type", 
     main = "Household types per age category",
     col=c("steelblue1","sienna1","plum2","goldenrod1"), las = 1, mar = c(3,4,4,2))

# Perform chi square test to evaluate whether age and household type are independent
svychisq(~a_agegr11_dv + hh_type, design=Design, statistic = "Chisq")

# Remove this part??? #######################################################################
# >>>>>>> I think this is more interesting with the 5-year variable?
tab <- svytable(~a_agegr11_dv + hh_type, Design) # creates a contingency table  
# The Ntotal argument can be either a single number or a data frame whose first 
# column gives the (first-stage) sampling strata and second column the population size 
# in each stratum. 
# In this second case the svytable command performs ???post-stratification???: 
# tabulating and scaling to the population within strata and then adding up the strata.
ftable(tab)
summary(ftable(tab))

######################################################################## 
```
# Question 6
```{r}
# Summarise a_ivfio
(summary <- summary(society$a_ivfio))
summary[summary != 0]
adults <- sum(society$a_ivfio == "Full interview") #the full interviews with adults
youths <- sum(society$a_ivfio ==  "Youth Interview") #the interviews with children
sum(adults, youths) #all personally completed interviews

# The nonresponse indicator is 1 for all (partial) nonresponse and 0 for full 
# (youth) interviews
society$NR <- 1
society$NR[society$a_ivfio == "Youth Interview"] <- 0
society$NR[society$a_ivfio == "Full interview"] <- 0
```


# Question 7
```{r}
# Further investigate the level Language difficulties
society$a_iproxy[society$a_ivfio=="Language difficulties"]
# It's clearly a complete nonresponse

# Repeat creation of model design from Q4 so NR variable is included
Design <- svydesign(ids=~a_hidp, strata=~a_strata, data=society, weights=~a_psnenip_xw)
Nonmiss <- with(Design, subset(Design, a_employ!="refuse" & a_employ!="don\'t know"))
# Create model
I_personal <- with(Nonmiss, subset(Nonmiss, NR==0))
                   
# Calculate nonresponse
svymean(~employ_dv, I_personal)
confint(svymean(~employ_dv, I_personal))

# With NR: 46,2% of the population is employed, with a 95% confidence interval of 44.3%-48.1%
# Exclude NR: 50.4% of the population is employed, with a 95% confidence interval of 48.2%-52.6%

# May be caused by children of younger than 10, which were not interviewed 
# (and thus nonresponders) and are not employed 
summary(society$employ_dv[society$a_dvage < 10])
summary(factor(society$a_ivfio[society$a_dvage < 10]))

# Create design that excludes children under the age of ten:
Nochild <- with(Nonmiss, subset(Nonmiss, a_dvage >= 10))

# Calculate nonresponse
svymean(~employ_dv, Nochild)
confint(svymean(~employ_dv, Nochild))
# Estimated proportion 52.0%, 95% CI[50.0%, 54.1%]
```

# Question 8
```{r}
# Investigate the nonresponse
mean(society$NR) # compute proportion nonresponders
# Look at levels of household response
summary(society$a_hhresp_dv) 
# All households at least filled in the grid, so we only look into person (unit) nonresponse.
# use design weights, 
#NRdesign <- svydesign(ids=~a_hidp, strata=~a_strata, data=society, weights=~ )

##### Remove and change problematic variables
# Combine races
society$a_racel_dv <- factor(society$a_racel_dv)
levels(society$a_racel_dv) <- c(
  rep("missing", 4), "UK Native", rep("White - Nonnative", 2), rep("Mixed", 4), 
  rep("Asian or Asian British", 4), rep("Black/African/Carribean", 3), rep("Other",2))

# Change livewith of same-sex couples to 'yes'
society$a_livewith <- factor(society$a_livewith)
levels(society$a_livewith) <- c("inapplicable", "yes", "no", "yes")

# Combine respm16_dv and respf16_dv
society$a_resp16_dv <- "No"
society$a_resp16_dv[society$a_respf16_dv == "Yes" | society$a_respf16_dv == "Yes"] <- "Yes" 



##### MODELS
# model with all predictors
# fullmodel <- glm(NR ~ a_gor_dv + a_urban_dv + a_sex + a_racel_dv + 
#                    a_employ + a_dvage + a_agegr5_dv + a_agegr10_dv + a_agegr13_dv + 
#                    a_livesp_dv +  a_livewith + a_cohab_dv + #a_mastat_dv +
#                    a_single_dv + a_depchl_dv + a_rach16_dv + a_resp16_dv + 
#                    a_nchild_dv + hh_type + hh_size + n_child, family = binomial, data = society)
# 
# fullmodel$coefficients
# Doesn't produce errors

# Create new model without variables that constitute singularity errors:
halffullmodel <- glm(NR ~ a_gor_dv + a_urban_dv + a_sex + a_dvage + 
                       a_agegr5_dv + a_agegr10_dv + a_agegr13_dv + hh_size + n_child + 
                       hh_type, family = binomial, data = society)
# we delete a_racel_dv and a_employ because it is only a good predictor because it includes 'missing'

# model without any predictors
emptymodel <- nothing <- glm(NR ~ 1, family=binomial, data = society )

# backwards selection
# backwards <- step(fullmodel)
halfbackwards <- step(halffullmodel)

# forwards selection to compare backwards model with it. 
# forwards <- step(emptymodel,
# scope=list(lower=formula(emptymodel),upper=formula(fullmodel)), direction="forward")

halfforwards <- step(emptymodel,scope=list(lower=formula(emptymodel),
                                           upper=formula(halffullmodel)), direction="forward")

# show different models: 
# formula(fullmodel)
# formula(backwards)
# formula(forwards)
formula(halfforwards)
formula(halffullmodel)
formula(halfbackwards)

# forwards and backwards logistic regression lead to the same model! 

# dive deeper into halfbackwards since this is our model of interest
summary(halfbackwards)

# evaluate backwards model compared with backwards regression/full model  
anova(halfbackwards, emptymodel, test = "Chisq")  # backwards model does a better job at prediction than empty model. 
anova(halfbackwards, halffullmodel, test = "Chisq") # backwards model does an equally good job at prediction as the full model. 

# so we now belief that the following variables are predictors of non-response: 
# a_racel_dv + a_agegr5_dv + a_agegr13_dv + hh_size + a_sex + 
#     a_gor_dv + hh_type + a_urban_dv
summary(halfbackwards)
# significant variables are a_dvage, hh_size, hh_type, a_sex, a_urban_dv and a_gor_dv, n_child. 
# significant with p < .001 B's come from variables a_sex, hh_size and hh_type (couple without children)

# just some exploration of response distribution of individual variables 
# plot nonresponse 
tbl <- table(society$NR)
#t1 <- tbl/3600*100 
tbl <- prop.table(tbl, margin = 1)

png("nonresponse.png")
plot(tbl, ylab = "nonresponse = 1", 
     main = "distribution of nonresponse",
     col=c("steelblue1","sienna1"), las = 1)
dev.off()
# plot relationship sex / nonresponse 

t2 <- table(as.matrix(society$a_sex),as.matrix(society$NR)) 
t2 <- prop.table(t2, margin = 1)
# females respond more often then males do 
png("sex.png")
plot(t2, xlab = "sex", ylab = "nonresponse = 1", 
     main = "distribution of nonresponse",
     col=c("steelblue1","sienna1"), las = 1)
dev.off()
# plot relationship hh_size / non response
t4 <- table(society$hh_size, society$NR) 
t4 <- prop.table(t4, margin =1)

# the bigger the household, the more nonresponders 
png("hh_size.png")
plot(t4, xlab = "household size", ylab = "nonresponse = 1", 
     main = "distribution of nonresponse",
     col=c("steelblue1","sienna1"), las = 1)
dev.off()

# plot relationship hh_type / non response 
t5 <- table(society$hh_type, society$NR)
t5 <- prop.table(t5, margin = 1)
png("hh_type.png")
plot(t5, xlab = "household type", ylab = "nonresponse = 1", 
     main = "distribution of nonresponse",
     col=c("steelblue1","sienna1"), las = 1)
dev.off()


# plot relationship age / nonresponse 
t3 <- table(society$a_agegr13_dv, society$NR)
t3 <- prop.table(t3, margin =1)
plot(t3, xlab = "age", ylab = "nonresponse = 1", 
     main = "distribution of nonresponse",
     col=c("steelblue1","sienna1"), las = 1) # the higher the age category, the less nonresponse.


```

